---
title: "Steak Preferences in the United States"
subtitle: "Summary Report For the _**`r params$region`**_ Region"
author: "Author: Irene Espinosa-Martos, PhD."
date: "`r format(Sys.Date(), '%d de %B, %Y')`"
output: html_document
params:
  region:
    label: "Region:"
    value: Mountain
    input: select
    choices: ["East North Central" ,
     "South Atlantic" , "New England" ,
     "Middle Atlantic" , "West South Central" ,
     "West North Central" , "Pacific" ,           
     "Mountain" , "East South Central" ]
---

```{r setup, include=FALSE}
# define knitr options
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message =  FALSE)

# set any missing NAs in tables to blank
options (knitr.kable.NA = '')

# install and/or load R packages
libraries <- c("fivethirtyeight", "tidyverse", "knitr" ,"kableExtra", "formattable" , "ggthemes")
check.libraries <- is.element(libraries, installed.packages()[, 1])==FALSE
libraries.to.install <- libraries[check.libraries]
if (length(libraries.to.install!=0)) {
  install.packages(libraries.to.install)
}
success <- sapply(libraries,require, quietly = FALSE,  character.only = TRUE)
if(length(success) != length(libraries)) {stop("A package failed to return a success in require() function.")}

# load steak_survey dataset
data("steak_survey", package="fivethirtyeight")

# create a subset, no missing NAs pick a specific region
sdat <- na.omit(steak_survey) %>%
  filter (region==params$region)

```

\pagebreak

## Background

In May 2014, Walt Hichey at [538.com](http:77fivethirtyeight.com/) published an article entitled ["How Americans Like Their Steak"](https://fivethirtyeight.com/features7how-americans-like-their-steak/). This article utilized survey data collected from 550 people which asked questions related to various risky activities, such as whether they speed in traffic, if they smoke or go skydiving, and if they prefer a riskier lottery. The survey also asked if they eat steak and if so, how they prefer their steak.

## Purpose

The "steak survey" dataset that was used for the ["How Americans Like Their Steak"](https://fivethirtyeight.com/features7how-americans-like-their-steak/) article, is available through the R package [`fivethirtyeight`](https://cran.r-project.org/web/packages/fivethirtyeight/index.html). Using this dataset, this article summarize the steak preferences for the _**`r params$region`**_ region of the US. The data subset used for our summary had `r nrow(sdat)` respondents.

## Demographic Summary Tables

The age categories of the `r nrow(sdat)` survey respondents were:

```{r}
# create table of the age categories
tb <- sdat %>%
  select(age) %>%
  table() %>%
  prop.table() * 100

# convert table to data frame
dt <- as.data.frame(tb)

# use kable from knitr package to make table and use kable styling from kableExtra package
knitr::kable(dt, format = "html",
             col.names = c("Ages", "%"),
             digits = 2,
             caption = "Ages of Survey Respondents") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

The education levels of the `r nrow(sdat)` survey respondents were:

```{r}
# repeat for education
tb <- sdat %>%
  select(educ) %>%
  table() %>%
  prop.table() * 100

# convert table to data frame
dt <- as.data.frame(tb)

# use kable from knitr package to make table and use kable styling from kableExtra package
knitr::kable(dt, format = "html",
             col.names = c("Education", "%"),
             digits = 2,
             caption = "Education of Survey Respondents") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

The income levels of the `r nrow(sdat)` survey respondents were:

```{r}
# repeat for income
tb <- sdat %>%
  select(hhold_income) %>%
  table() %>%
  prop.table() * 100

# convert table to data frame
dt <- as.data.frame(tb)[-6,]

# use kable from knitr package to make table and use kable styling from kableExtra package
knitr::kable(dt, format = "html",
             col.names = c("Income", "%"),
             digits = 2,
             caption = "Income of Survey Respondents") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

## _OPTIONAL_ Demographics of Survey Respondents in a Merged Table

This section is provided as an example of how to merge the three demographic summary tables for the `r nrow(sdat)` survey respondents above into a single merged table.

```{r}
# create table summary for ages
tb1 <- sdat %>%
  select(age) %>%
  table() %>%
  prop.table() * 100
# create table summary for education
tb2 <- sdat %>%
  select(educ) %>%
  table() %>%
  prop.table() * 100
# create table summary for income
tb3 <- sdat %>%
  select(hhold_income) %>%
  table() %>%
  prop.table() * 100

# merge and convert tables to data frame
dt1 <- as.data.frame(tb1)
dt2 <- as.data.frame(tb2)
dt3 <- as.data.frame(tb3)[-6,]
dt <- merge(dt1,dt2, by = 0, all= TRUE)[-1]
dt <- merge (dt, dt3,  by = 0, all = TRUE)[-1]

# use kable from knitr package to make table and use kable styling from kableExtra package
knitr::kable(dt, format = "html",
             col.names = c("Category", "%","Category", "%","Category", "%"),
             digits = 2,
             caption = "Demographics of Survey Respondents") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE) %>%
            kableExtra::add_header_above( c( "Age" = 2, "Education" = 2, "Income" = 2))

```

## Steak Preparation Preference by Gender

Finally, here is a breakdown of the `r nrow(sdat)` survey respondents for the _**`r params$region`**_ region of the Us on how they prefer their steak to be prepared by gender.


```{r}
ggplot(sdat,
       aes(x = steak_prep, fill = female)) +
    geom_bar(position = "dodge", colour="black") +
    ggtitle(paste0("Steak Preference by Gender: ",
             params$region," Region")) +
    xlab("Steak Preparation Preference") + 
    ylab("Number of Respondents") +
    scale_fill_manual(values=c("skyblue", "palevioletred"),
      name="Gender", 
      breaks=c(FALSE,TRUE),
      labels=c("Male", "Female")) +
    theme_fivethirtyeight()

```




Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
