---
title: "Weather Information Preferences in the United States"
subtitle: "Summary Report For the _**`r params$region`**_ Region"
author: "Author: Irene Espinosa-Martos, PhD."
date: "`r format(Sys.Date(), '%d de %B, %Y')`"
output: html_document
params:
  region:
    label: "Region:"
    value: Mountain
    input: select
    choices: ["East North Central" ,
     "South Atlantic" , "New England" ,
     "Middle Atlantic" , "West South Central" ,
     "West North Central" , "Pacific" ,           
     "Mountain" , "East South Central" ]
---

```{r setup, include=FALSE}
# define knitr options
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message =  FALSE)

# set any missing NAs in tables to blank
options (knitr.kable.NA = '')

# install and/or load R packages
libraries <- c("fivethirtyeight", "tidyverse", "knitr" ,"kableExtra", "formattable" , "ggthemes", "stringr")
check.libraries <- is.element(libraries, installed.packages()[, 1])==FALSE
libraries.to.install <- libraries[check.libraries]
if (length(libraries.to.install!=0)) {
  install.packages(libraries.to.install)
}
success <- sapply(libraries,require, quietly = FALSE,  character.only = TRUE)
if(length(success) != length(libraries)) {stop("A package failed to return a success in require() function.")}

# load steak_survey dataset
data("weather_check", package="fivethirtyeight")

# create a subset, no missing NAs pick a specific region
sdat <- weather_check %>%
  filter (region==params$region)

```

\pagebreak

## Background

In April 2015, Walt Hichey at [538.com](http:77fivethirtyeight.com/) published an article entitled ["Where People Go To Check the Weather"](https://fivethirtyeight.com/features/weather-forecast-news-app-habits/). This article utilized survey data collected from 928 people which asked questions on if they typically check the weather on a daily basis and if so, what source they used to check the weather. Additional follow-up questions were asked about which specific websites or apps they used an ifn they had a smartwatch if they would check the weather on that smartwatch.

## Purpose

The "steak survey" dataset that was used for the ["Where People Go To Check the Weather"](https://fivethirtyeight.com/features/weather-forecast-news-app-habits/) article, is available through the R package [`fivethirtyeight`](https://cran.r-project.org/web/packages/fivethirtyeight/index.html). Using this dataset, this article summarizes the weather source preferences for the _**`r params$region`**_ region of the United States by gender. The data subset used for our summary had `r nrow(sdat)` respondents.

## Demographics of Survey Respondents in a Merged Table

This section is provided as an example of how to merge the three demographic summary tables for the `r nrow(sdat)` survey respondents above into a single merged table.

```{r}
# create table summary for ages
tb1 <- sdat %>%
  select(age) %>%
  table() %>%
  prop.table() * 100
# create table summary for gender
tb2 <- sdat %>%
  select(female) %>%
  table() %>%
  prop.table() * 100
# create table summary for income
tb3 <- sdat %>%
  select(hhold_income) %>%
  table() %>%
  prop.table() * 100

# merge and convert tables to data frame
dt1 <- as.data.frame(tb1)
dt2 <- as.data.frame(tb2)
dt3 <- as.data.frame(tb3)[-6,]
dt <- merge(dt1,dt2, by = 0, all= TRUE)[-1]
dt <- merge (dt, dt3,  by = 0, all = TRUE)[-1]

# use kable from knitr package to make table and use kable styling from kableExtra package
knitr::kable(dt, format = "html",
             col.names = c("Category", "%","Category", "%","Category", "%"),
             digits = 2,
             caption = "Demographics of Survey Respondents") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE) %>%
            kableExtra::add_header_above( c( "Age" = 2, "Gender" = 2, "Income" = 2))

```

## Steak Preparation Preference by Gender

Finally, here is a breakdown of the `r nrow(sdat)` survey respondents for the _**`r params$region`**_ region of the Us on how they prefer their steak to be prepared by gender.


```{r}
ggplot(sdat,
       aes(x = weather_source, fill = female)) +
    geom_bar(position = "dodge", colour="black") +
    ggtitle(paste0("Weather source by Gender: ",
             params$region," Region")) +
    xlab("Weather Source Preference") + 
    ylab("Number of Respondents") +
    scale_fill_manual(values=c("skyblue", "palevioletred"),
      name="Gender", 
      breaks=c(FALSE,TRUE),
      labels=c("Male", "Female")) +
    coord_flip() +
    theme_fivethirtyeight()

```




Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
